In conclusion, this study presents an efficient algorithm for updating PageRank on dynamic graphs. Given a batch update of edge insertions and deletions, our improved Dynamic Frontier (DF) and Dynamic Frontier with Pruning (DF-P) approaches identify an initial set of affected vertices and incrementally expand, and optionally contract/prune (with DF-P PageRank) this set across iterations. We observe that, expanding the frontier based on relative change in rank $\Delta r/r$ with a frontier tolerance $\tau_f$ of $10^{-6}$, and a corresponding prune tolerance $\tau_p$ of $10^{-6}$ (for DF-P PageRank) yields the best performance, while achieving lower error rates than Static PageRank.

On a server equipped with a 64-core AMD EPYC-7742 processor, DF PageRank demonstrates average speedups of $8.0\times$, $4.5\times$, and $3.2\times$ compared to Static PageRank when processing real-world dynamic graphs with batch updates of sizes $10^{-5}|E_T|$, $10^{-4}|E_T|$, and $10^{-3}|E_T|$, respectively. Additionally, it surpasses Dynamic Traversal (DT) PageRank, a commonly used method for updating PageRank on dynamic graphs, by $1.3\times$, $1.1\times$, and $1.5\times$ for the same batch updates. DF-P PageRank achieves even higher speedups, averaging $26.2\times$, $11.9\times$, and $7.5\times$ over Static PageRank, and $4.2\times$, $2.8\times$, and $3.6\times$ over DT PageRank for identical batch updates. For real-world dynamic graphs, we recommend DF-P PageRank, with a suggestion to switch to DF PageRank if higher error is observed.

For batch updates ranging from $10^{-7}|E|$ to $10^{-3}|E|$ with $80\%$ insertions and $20\%$ deletions on large static graphs, DF PageRank demonstrates average speedups of $7.2\times$, $2.6\times$, and $4.0\times$ compared to Static, ND, and DT PageRank respectively. Meanwhile, DF-P PageRank achieves average speedups of $9.6\times$, $3.9\times$, and $5.6\times$ over the same approaches. For large graphs with random updates, we recommend DF-P PageRank, except for web graphs, where we suggest selecting DF PageRank.

Using $16$ threads, DF PageRank exhibits an average speedup of $11.5\times$ compared to single-threaded execution, indicating a performance boost of $1.8\times$ for each doubling of threads. Conversely, DF-P PageRank achieves an average speedup of $8.8\times$, implying a performance increase of $1.7\times$ for each doubling of threads.
